---
title: "h1a_h1c_analysis"
output: html_document
date: "2024-11-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview of Hypotheses

## H1

Consistent with our prior research (Lin, Adolphs, & Alvarez, 2018), we expect to replicate that corruptible looks predict corruption records using colored photos of the same set of politicians as in our prior research and ratings from U.S. representative participants.

### H1a

We expect that participants perceive politicians with corruption records to be more corruptible than politicians with clean records, and vice versa.

### H1b

We expect that the association between corruptible looks and corruption records will remain robust when controlling for other features of the photos that may influence corruptibility perception.

### H1c

We expect that the association between face perception and corruption records will be significant for corruptibility-related traits (e.g., corruptible, genuine) but not for corruptibility-unrelated traits (e.g., attractiveness).

## Focus

This analysis report will include the analysis for **H1a** and **parts of H1c**. Since H1c builds upon H1a by extending the t-test analysis to related (Genuine) and unrelated traits (Attractive & Dominant), we will include the relevant analysis and tests of H1c together with H1a in this report to avoid redundancy.

------------------------------------------------------------------------

# Testings for H1a

To test H1a, we will conduct an aggregate-level t-test for the rating of corruptibility. The cleaned data will include the average rating of all four traits for a given politician across all participants. We will test whether politicians with corruption records are rated more corruptible on average than politicians with clean records using a two-sample one-sided t-test.

```{r}
#loading required library

library(tidyverse)
library(dplyr)
```

```{r}
#Load the clean data set

dat <- read.csv("cleaned.csv")
```

# Dataset Overview

The dataset `cleaned.csv` contains information about politicians and their ratings on various traits. Each row represents a politician, and the columns include:

-   **`convicted`**: Binary variable indicating whether the politician has a corruption record:
    -   `1`: Politician has a corruption record.
    -   `0`: Politician has no corruption record.
-   **`Corruptible`**, **`Genuine`**, **`Attractive`**, **`Dominant`**: Numeric values reflecting the average trait ratings for each politician across every participants.
-   The rest of the variables are not applicable for H1a analysis.

------------------------------------------------------------------------

## Summary of the Dataset

```{r dataset-summary}
# Display the first few rows of the dataset
head(dat)

# Summarize the dataset structure
summary(dat)
```

## Function Definition: **`analyze_trait`**

We define a function `analyze_trait` to:

-   Compute t-tests for specified traits.
-   Dynamically handle the direction of the alternative hypothesis:
    -   For traits **Corruptibility** and **Dominant**, we test whether $x$ (corrupted) \> $y$ (clean) (the alternative hypothesis will be `"greater"`).
    -   For traits **Genuine** and **Attractive**, we test whether $y$ (clean) \> $x$ (corrupted) (the alternative hypothesis will be `"less"`).

```{r}
analyze_trait <- function(trait_type) {
  
  # Step 1: Create explicit groups for the t-test
  clean <- dat %>% filter(convicted == 0) %>% pull(trait_type)
  corrupted <- dat %>% filter(convicted == 1) %>% pull(trait_type)
  
  # Step 2: Dynamically determine the alternative hypothesis for different traits
  alternative_hypothesis <- ifelse(trait_type %in% c("Genuine", "Attractive"), "less", "greater")
  
  # Step 3: Perform the t-test
  t_test_result <- t.test(
    x = corrupted,  # Ratings for the corruption group
    y = clean,      # Ratings for the clean group
    alternative = alternative_hypothesis  # Dynamic alternative
  )
  
  # Step 4: Output results
  cat("\n===== T-TEST RESULTS FOR TRAIT:", trait_type, "=====\n")
  print(t_test_result)
}

```

# H1a: Corruptibility

```{r}
corruptible_t_test <- analyze_trait("Corruptible")
```

# H1c: Testing for Other Traits

## Genuine

```{r}
genuine_t_test <- analyze_trait("Genuine")
```

## Attractive

```{r}
attractive_t_test <- analyze_trait("Attractive")

## Dominant

dominant_t_test <- analyze_trait("Dominant")
```
