---
title: "h1b_h1c_analysis"
output: html_document
date: "2024-11-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview of Hypotheses

## H1

Consistent with our prior research (Lin, Adolphs, & Alvarez, 2018), we expect to replicate that corruptible looks predict corruption records using colored photos of the same set of politicians as in our prior research and ratings from U.S. representative participants.

### H1a

We expect that participants perceive politicians with corruption records to be more corruptible than politicians with clean records, and vice versa.

### H1b

We expect that the association between corruptible looks and corruption records will remain robust when controlling for other features of the photos that may influence corruptibility perception.

### H1c

We expect that the association between face perception and corruption records will be significant for corruptibility-related traits (e.g., corruptible, genuine) but not for corruptibility-unrelated traits (e.g., attractiveness).

## Focus

This analysis report will include the analysis for **H1b** and **parts of H1c** Since H1c builds upon H1b by extending the logistic regression analysis to related (Genuine) and unrelated traits (Attractive & Dominant), we will include the relevant analysis and tests of H1c together with H1b in this report to avoid redundancy.

------------------------------------------------------------------------

# Testings for H1b

To test H1b, we will conduct an aggregate-level logistic regression. The cleaned data will include the average rating of all four traits for a given politician across all participants. We will regress the politicians’ records (corrupt or clean) on their average corruptibility ratings while controlling for photo characteristics that may influence corruptibility perception such as the politician’s age, image clarity, image sources, smile intensity, with or without glasses and jewelry, and hair and clothing colors, etc.

```{r}
#Loading library for regression model 
library(minqa)
library(nloptr)
library(RcppEigen)
library(lme4)
```

```{r}
#Load the clean data set

dat <- read.csv("cleaned.csv")
```

# Dataset Overview

The dataset `cleaned.csv` contains information about politicians and their ratings on various traits. Each row represents a politician, and the columns include:

-   **`convicted`**: Binary variable indicating whether the politician has a corruption record:
    -   `1`: Politician has a corruption record.
    -   `0`: Politician has no corruption record.
-   **`Corruptible`**, **`Genuine`**, **`Attractive`**, **`Dominant`**: Numeric values reflecting the average trait ratings for each politician across every participants.
-   The rest of the variables are not applicable for H1a analysis.

------------------------------------------------------------------------

## Summary of the Dataset

```{r dataset-summary}
# Display the first few rows of the dataset
head(dat)

# Summarize the dataset structure
summary(dat)
```

## Function Definition: **`run_regression`**

We define a function `run_regression` to:

-   Compute t-tests for specified traits.
-   Dynamically handle the trait type for the regression formula

```{r}
run_regression <- function(trait) {
  
  # Dynamically add the trait column to the model
  formula <- as.formula(paste("convicted ~", trait, "+ wiki + gov + news + age + glasses + beard + mustache + 
                               hair_blond + hair_white + hair_black + hair_brown + pixels + AU06 + AU12 + 
                               clothing_r + clothing_g + clothing_b"))
  
  # Fit the logistic regression model
  model <- glm(formula, data = dat, family = "binomial")

  # Return the model summary
  cat("\n===== MODEL SUMMARY FOR TRAIT:", trait, "=====\n")
  print(summary(model))
}
```

# H1b: Regression for Corruptibility

```{r}
corruptible_model <- run_regression("Corruptible")
```

# H1c: Regression for Other Traits

## Genuine

```{r}
genuine_result <- run_regression("Genuine")
```

## Attractive

```{r}
attractive_result <- run_regression("Attractive")
```

## Dominant

```{r}
dominant_result <- run_regression("Dominant")
```
